### 2025-06-25 lecturenotes_aima203/ch4_2_aima203.tex
chapter 4

\subsubsection{Drawback}
The linear splines derived above are continuous in $[x_0, x_n]$, but their slpes are discontinuous, i.e their first derivatives are discontinuous.

@@

\section{Cubic B splines}
The cubic spline formulae derived in the preceding section are global in nature, which means that they do not permit any local changes in the given data. The B-splines are the basis functions of the splines that facilitates us the local changes. \\
A B-spline of order $n$, is denoted by $s_{n,i}(x)$ is a spline of degree $(n-1)$ with knots $k_{i-n}, k_{i-n+1},...,k_i$, which is zero everywhere except in the interval $[k_{i-n}, k_i]$.
\begin{definition}
  A cubic B-spline $s_{4,i}(x)$ is a spline of degree $3$ with knots $k_{i-4}, k_{i-3},k_{i-2},k_{i-1},k_i$, such that
  \begin{enumerate}
  \item on each interval, $s_{4,i}(x)$ is a polynomial of degree $3$ or less,
  \item $s_{4,i}(x), \; s'_{4,i}(x), \; s''_{4,i}(x)$ are continuous, and
   \item $s_{4,i}(x) > 0$ inside $[k_{i-4}, k_i]$.
  \end{enumerate}
\end{definition}
\end{document}

>>>>

\subsubsection{Drawback}
The linear splines derived above are continuous in $[x_0, x_n]$, but their slopes are discontinuous, i.e their first derivatives are discontinuous.

@@

\subsection{Exercise}
\begin{enumerate}
\item For the data points: $(0,0), \; (\pi /2,1), \; (\pi,0)$, determine the following:
  \begin{enumerate}
  \item natural quadratic splines
  \item natural cubic splines
  \item $y(\pi /6)$ using natural cubic spline
  \end{enumerate}
\item Determine $y(\pi/6)$ using the natural cubic splines from the data points: \\ $(0,0), \; (\pi /4,1/ \sqrt{2}), \; (\pi /2 ,1), \; (3\pi /4, 1/ \sqrt{2}), \; (\pi,0)$.

\item What do you understand by natural spline? Explain why the natural cubic spline condition have $M_0 = 0$ and $M_n = 0$.
  \end{enumerate}

\end{document}
